ZCB Pricing

### 2018-07-27 ZCB Pricing .Rmd
MC Comments and adapting finite difference methods

bond_price <- function(alpha=0.08,beta=0.8,sigma=0.005,t=0,T=1,r=0.08){
  A <- (1-exp(-beta*(T-t)))/(beta)
  B <- (1/beta)*((alpha-(sigma)^2/(2*beta))*(A - T +t) - (sigma^(2)*A^2)/(4))
  exp(-A*r +B)
}

@@

```{r 2}
#################################################################################
# Crude Monte Carlo
#################################################################################
Crude_Monte_Carlo=function(alpha =0.08,beta=0.8,sigma=0.005,T=1,r0 = 0.08,n=1000,s=100){

  #Simulate interest rates
  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path[1,]= r0

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*rnorm(1)# Euler Discretization
      sim_path[j,i]=sim_path[j-1,i]+dr
    }
  }

@@

#################################################################################
# Antithetic Variates
################################################################################

Antithetic_variates= function(alpha =0.08,beta=0.8,sigma=0.005,T=1,r0 = 0.08,n=1000,s=100){
  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path_inv= matrix(0,nrow = s+1, ncol = n)# matrix to generate interest paths using inverse RV
  sim_path[1,]= r0
  sim_path_inv[1,]=r0
  z = rnorm(n)
  z. = -z

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z[i] # Euler Discretization
      dr_i= (alpha-beta*sim_path_inv[j-1,i])*dt + sigma*sqrt(dt)*z.[i]
      sim_path[j,i]=sim_path[j-1,i]+dr
      sim_path_inv[j,i]=sim_path_inv[j-1,i]+dr_i

@@

  # Monte Carlo Simulations
  Sum=colSums(sim_path[2:(s+1),]*dt)
  Sum.=colSums(sim_path_inv[2:(s+1),]*dt)
  int.=exp(-Sum.)
  int=exp(-Sum)# Estimate using the conjugate rates
  Average=(int+int.)/2 # Average of estimates
  mean=mean(Average)
  sd=sd(Average)/sqrt(n)
  result= mean
  result

@@

Moment_Matching=function(alpha =0.08,beta=0.8,sigma=0.005,T=1,r0 = 0.08,n=1000,s=100){

@@

# Simple Bond price function
B <- function(t,T,r){
  exp(-(T-t)*r)
}

explicit <- function(alpha= 0.08,beta=0.8,sigma=0.005,r=0.08,I=1:100,T=2){

# I = partition of time
# J = partition of rate
# dr = length of rate sub-intervals
# r must be in the middle of partition of J
# length of J must be twice the length of I
dr <- 2*r/(2*length(I))

J <- seq(0,2*r,dr)

dt = T/length(I) # length of each time sub-interval

V <- matrix(NA, nrow = length(J), ncol = length(I)+1) # empty matrix
nr <- length(J) # number of rows of the matrix - represents partition of rate
nt <- length(I)+1 # number of columns - represents partition of time

V[,length(I)+1] <- rep(1) # Boundary condition at expiry of bond

count <- 0 # set up count variable
for (i in rev(I)){ # rev is the reverse order of the time partition | i corresponds to the columns of the lattice
  count <- count+1 # keeping track of loop
  for (j in (1+count):(nr-count)){ # j corresponds to the rows of the lattice
    pu <- dt*(sigma^2/(2*dr^2) + (alpha -beta*J[nr-j])/(2*dr)) # up value
    pm <- 1 - (sigma^2*dt)/(dr^2) # middle value
    pd <- dt*(sigma^2/(2*(dr^2)) - (alpha -beta*J[nr-j])/(2*dr)) # down value
    V[j,i] <- ((1+J[nr-j]*dt)^(-1))*(pu*V[j-1,i+1] + pm*V[j,i+1] +pd*V[j+1,i+1])
  }

@@

V[(nr+1)/2,1]

#============================================================================================
#Doing some plots
# 
# xvec <- seq(-0.3,0.5,0.01)
# H <- NULL
# for (r in 1:length(xvec)){
#   S[r] <- bond_price(0.08,0.8,0.005,0,1,xvec[r])
# }
# 
# for (i in 1:length(xvec)){
#   H[i] <- explicit(r = xvec[i], I= 1:50, T=5)
# }
# 
# plot(H, x = xvec,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Crank - Nicolson vs Known Solution")
# lines(xvec,S, col ="blue", lty =1.5)

@@

  for (i in I){
    C <- matrix(0,ncol = 1, nrow = nr-1) # Another matrix for boundary conditions
    C[nr-1,1] <- (-pu(nr-1))*B[nr-1,1]
    C[1,1] <- -pd(1)
    A <- B + C # Working backwards recursively 
    B <- solve(Diag)%*%A 
  }

  B # Output of bond prices at time 0 for varying rates

@@

xvec <- seq(-0.4,0.6,0.005)
R <- implicit(T=5,J= xvec, I = 1:1250)
S <- NULL
for (r in 1:length(xvec)){
  S[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
  }
plot(R, x = xvec[1:length(xvec)-1],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Implicit Method vs Known Solution")
lines(xvec,S, col ="blue", lty =1.5)

@@

  Bond_values

@@

xvec <- seq(-0.4,0.8,0.005)
S <- NULL
for (r in 1:length(xvec)){
  S[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
}
plot(CN(), x = xvec[1:length(xvec)],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Crank - Nicolson vs Known Solution")
lines(xvec,S, col ="blue", lty =1.5)
```

>>>>

bond_price <- function(alpha=0.08,beta=0.8,sigma=0.005,t=0,T=5,r=0.08){
  A <- (1-exp(-beta*(T-t)))/(beta)
  B <- (1/beta)*((alpha-(sigma)^2/(2*beta))*(A - T +t) - (sigma^(2)*A^2)/(4))
  exp(-A*r +B)
}

@@

```{r 2}

# Crude Monte Carlo

Crude_Monte_Carlo=function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){

  # alpha, beta, sigma <- Fixed Parameters
  # T <- maturity time
  # r <- Initial rate at time 0
  # n <- Number of Simulations
  # s <- NUmber of Sub-intervals over time

  #Simulate interest rates
  dt <- T/s # Length of sub-intervals
  sim_path = matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path[1,]= r # Initial rates on first row.

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*rnorm(1)# Discretize the SDE of the interest model using Euler Discretization to simulate rates over discrete sub-intervals
      sim_path[j,i]=sim_path[j-1,i]+dr
    }
  }

  # Monte Carlo Simulations
  Sum = colSums(sim_path[2:(s+1),]*dt) # Approximate the integral using summation
  int = exp(-Sum) # Take the exponent of the Integral
  mean = mean(int) # Find the average of expression
  sd = sd(int)/sqrt(n)
  result = mean
  result
}

@@

# Antithetic Variates

Antithetic_variates = function(alpha =0.08,beta=0.8,sigma=0.005,T=1,r = 0.08,n=1000,s=100){
  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path_inv= matrix(0,nrow = s+1, ncol = n)# matrix to generate interest paths using negative values of RV
  sim_path[1,] = r
  sim_path_inv[1,] = r
  z = rnorm(n) # Generate standard normal values
  z. = -z # Estimate negative value

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z[i] # Same Euler Discretization as in the Crude Monte Carlo function
      dr_i= (alpha-beta*sim_path_inv[j-1,i])*dt + sigma*sqrt(dt)*z.[i] # Euler Discretization for negative rates
      sim_path[j,i]=sim_path[j-1,i]+dr
      sim_path_inv[j,i]=sim_path_inv[j-1,i]+dr_i

@@

  # Monte Carlo Simulations
  Sum=colSums(sim_path[2:(s+1),]*dt) # The Sum and Sum. variables record the summation of their respective rates to approximate the integral
  Sum.=colSums(sim_path_inv[2:(s+1),]*dt)
  int=exp(-Sum.) #Exponential of the estimate
  int.=exp(-Sum) #Exponential of the estimate under the negative rv 
  Average=(int+int.)/2   # The average of the expressions is taken for new estimate
  mean=mean(Average)  # The average of the expression is taken for the estimate
  sd=sd(Average)/sqrt(n)
  result= mean
  result

@@

Moment_Matching=function(alpha =0.08,beta=0.8,sigma=0.005,T=1,r0 = 0.08,n=100000,s=100){

@@

explicit = function(alpha= 0.08,beta=0.8,sigma=0.005,r=0.08,I=1:1250, J =seq(-0.4,0.6,0.005),T=5){

  # I = partition of time interval
  # J = partition of short rate interval

  dt <- T/N
  dr <- J[2] -J[1]

  pd <- function(j){
    ((J[j]+1/dt)^(-1))*(sigma^2/(2*(dr^2)) -(alpha -beta*J[j])/(2*dr))
  }

  pm <- function(j){
    ((J[j]+1/dt)^(-1))*(1/dt -sigma^2/(dr^2))
  }

  pu <- function(j){
    ((J[j]+1/dt)^(-1))*(sigma^2/(2*(dr^2)) + (alpha -beta*J[j])/(2*dr))
  }

  Diag_1 <- matrix(0,ncol = M, nrow = M)
    # Main diagonal of matrix 1

  for (j in 1:M){
    Diag_1[j,j] <- pm(j)

@@

  # Filling lower diagonal of matrix 1
  for (j in 2:M){
    Diag_1[j,j-1] <- pd(j)
  }

  # Filling upper diagonal of matirx 1
  for (j in 1:(M-1)){
    Diag_1[j,j+1] <- pu(j)
  }

  Bond_values <- matrix(1,nrow=M, ncol =1)

  for (i in I){
    Bond_values <- Diag_1%*%Bond_values
    Bond_values <- Bond_values
  }

  index = match(as.character(r),xvec)
  Bond_values[index]
}

#============================================================================================
#Doing some plots
# 
#  xvec <- seq(-0.4,0.6,0.005)
# S <- NULL
# for (i in 1:length(xvec)){
#    S[i] <- bond_price(alpha = 0.08,beta =0.8,sigma =0.005,t = 0, T = 5, r = xvec[i])
#  }
#  
# H <- explicit(r = 0.08, I= 1:100000, T=5,alpha = 0.08, beta =0.8, sigma= 0.005, J = xvec)
# 
#  
#  plot(H, x = xvec,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Explicit vs Known Solution")
#  lines(xvec,S, col ="blue", lty =1.5)

@@

  for (i in I){
    C <- matrix(0,ncol = 1, nrow = nr-1) # Another matrix for boundary conditions
    C[nr-1,1] <- (-pu(nr-1))*B[nr-1,1]
    C[1,1] <- -pd(1)
    A <- B  # Working backwards recursively 
    B <- solve(Diag)%*%A 
  }

  #B # Output of bond prices at time 0 for varying rates
  index = match(as.character(r),xvec)
  B[index]
}

@@

# xvec <- seq(-0.4,0.6,0.005)
# R <- implicit(T=5,J= xvec, I = 1:1250)
# S <- NULL
# for (r in 1:length(xvec)){
#   S[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
#   }
# plot(R, x = xvec[1:length(xvec)-1],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Implicit Method vs Known Solution")
# lines(xvec,S, col ="blue", lty =1.5)

@@

  #Bond_values

  index = match(as.character(r),xvec)
  Bond_values[index]
}

@@

# xvec <- seq(-0.4,0.8,0.005)
# S <- NULL
# for (r in 1:length(xvec)){
#   S[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
# }
# plot(CN(), x = xvec[1:length(xvec)],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Crank - Nicolson vs Known Solution")
# lines(xvec,S, col ="blue", lty =1.5)
```
